### 2021-12-19 Combinatorics/submission.tex
added introduction and conclusion to combinatorics

\documentclass{scrartcl}

\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{listings}
\usepackage{listings-rust}
\usepackage[style=english]{csquotes}
\usepackage[language=english, backend=biber, style=alphabetic, sorting=nyt]{biblatex}

\addbibresource{bibliography.bib}

\title{Miniproject - Combinatorics}
\author{Simon Pohmann}
\date{}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\powerset}{\mathfrak{P}}

\newcommand\restr[2]{{
    \left.\kern-\nulldelimiterspace
    #1
    \vphantom{\big|}
    \right|_{#2}
}}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{remark}[definition]{Remark}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{example}[definition]{Example}
\newtheorem{corollary}[definition]{Corollary}

\begin{document}
\maketitle

\subsection*{Introduction}
In this work, we will look at the LYM inequality
\begin{equation*}
    \sum_{i = 0}^n \frac {|A \cap L_i|} {|L_i|} \leq 1 \quad \text{for an antichain $A \subseteq \powerset(n)$}
\end{equation*}
and see how it generalizes to other posets than $\powerset(n)$.
In particular, it will be necessary to study posets that can be decomposed into layers $L_0, ..., L_n$, which we will call graded posets.
The goal then is to give a characterization of which posets satisfy such an inequality and then apply this to two examples.
These are the collection of partitions of $[n]$, ordered by refinement, and the collection of multi-subsets of a multiset.

We use the convention that $\N = \{ n \in \Z \ | \ n \geq 0 \}$.

\section{Part I}
First of all, we introduce the notion of a rank function on a graded poset.
\begin{proposition}
    \label{prop:definition_rank_function}
    Let $P$ be a graded poset in which every maximal chain has length $n + 1$. Then the function
    \begin{equation*}
        r: P \to \{ 0, ..., n \}, \quad x \mapsto \max\{ k \in \N \ | \ \exists a_1, ..., a_k \in P: \ a_1 < a_2 < ... < a_k < x \}
    \end{equation*}
    is well-defined and the unique function such that $x < y$ implies $r(x) < r(y)$ for all $x, y \in P$.
    We call it the \emph{rank function} of $P$.
\end{proposition}
\begin{proof}
    Clearly $r$ is well-defined, as for every increasing sequence $a_1 < ... < a_k < x$, we have the chain $C := \{ a_1, ..., a_k, x \}$ of size $k + 1$.
    Hence by assumption, $k \leq n$ and so $r(x)$ is finite and in $\{ 0, ..., n \}$.

    Further, consider $x < y$ in $P$.
    We have a sequence $a_1 < ... < a_{r(x)} < x$ by definition of $r$.
    It follows that there is an increasing sequence $a_1 < ... < a_{r(x)} < x < y$ and so $r(y) \geq r(k) + 1 > r(x)$.

    Lastly, assume there was another function $r': P \to \{ 0, ..., n \}$ with this property.
    Consider any $x \in P$.
    By definition of $r$, there is an increasing sequence $a_1 < ... < a_{r(x)} < x$ in $P$.
    Now consider a maximal chain $C$ containing the chain $\{ a_1, ..., a_{r(x)}, x \}$.

    Say $C = \{ b_1, ..., b_{n + 1} \}$ with $b_1 < ... < b_{n + 1}$ and $x = b_j$.
    Note that we have the increasing sequence $b_1 < ... < b_{j - 1} < b_j = x$ and so by the definition of $r$, find $j \leq r(x) + 1$.
    On the other hand, have $a_1, ..., a_{r(x)} \in C$ and thus $j = r(x) + 1$, i.e.
    \begin{equation*}
        a_1 = b_1, \ ..., \ a_{r(x)} = b_{r(x)}, \ x = b_{r(x) + 1}
    \end{equation*}
    As $b_i < b_{i + 1}$, we know that $r'(b_i) < r'(b_{i + 1})$ and inductively, we see $r'(b_i) \geq i - 1$.
    However, $r'(b_{n + 1}) \leq n$ and thus $r'(b_i) = i - 1$.
    Finally it follows that $r'(x) = r'(b_{r(x) + 1}) = r(x)$.
\end{proof}
Now we will show some basic properties of the rank function.
\begin{proposition}
    Let $P$ be a graded poset of maximal rank $n$ with rank function $r$.
    Then
    \begin{itemize}
        \item $x \in P$ is minimal iff $r(x) = 0$ and maximal iff $r(x) = n$.
        \item If $x < y$ and $r(x) + 1 \neq r(y)$ then there is $z \in P$ with $x < z < y$.
        \item If $x < y$ then there is an increasing sequence $x < a_1 < ... < a_{r(y) - r(x) - 1} < y$ in $P$. 
    \end{itemize}
\end{proposition}
\begin{proof}
    For (i), let $x \in P$ be minimal. Then there is no increasing sequence $a_1 < x$ in $P$, so $r(x) \leq 0$.
    Conversely, let $r(x) = 0$. Assume there was $y \in P$ with $y < x$, then $r(y) < r(x) = 0$, a contradiction.
    The analogous statement for maximal elements is proved in the same way.

    For (iii), consider $x < y$ in $P$. Then the chain $\{ x, y \}$ is contained in a maximal chain $C \subseteq P$.
    Say $C = \{ b_1, ..., b_{n + 1} \}$ with $b_1 < ... < b_{n + 1}$.

    Hence we find $r(b_i) < r(b_{i + 1})$ and so inductively that $r(b_i) \geq i - 1$ and $r(b_i) \leq i - 1$ since $r(b_1) \geq 0$ and $r(b_{n + 1}) \leq n$.
    It follows that $r(b_i) = i - 1$ and so $x = b_{r(x) + 1}, y = b_{r(y) + 1}$.
    Therefore, we have a chain
    \begin{equation*}
        x < b_{r(x) + 2} < ... < b_{r(y)} < y
    \end{equation*}
    of length $r(y) - r(x) - 1$.
    Statement (ii) follows directly, as in this setting, have $r(y) \geq r(x) + 2$ and so $r(y) - r(x) - 1 \neq 0$.
\end{proof}

\section{Part II}
In this part, we want to study whether graded posets satisfy an analogue of the LYM inequality.
\begin{proposition}
    \label{prop:characterization_lym}
    For a graded poset $P$ with layers $L_0, ..., L_n$ the following statements are equivalent:
    \begin{itemize}
        \item For every antichain $A \subseteq P$ have
        \begin{equation*}
            \sum_{i = 0}^n \frac {|A \cap L_i|} {|L_i|} \leq 1
        \end{equation*}
        \item For each $1 < i \leq n$ and $F \subseteq L_i$ have
        \begin{equation*}
            \frac {|\partial F|} {|L_{i - 1}|} \geq \frac {|F|} {|L_i|}
        \end{equation*}
        where
        \begin{equation*}
            \partial F := \{ a \in L_{i - 1} \ | \ \exists b \in F: \ a \leq b \}
        \end{equation*}
        \item There exists a (nonempty) sequence of maximal chains $C_1, ..., C_t$ such that for all $1 \leq i \leq n$ we have:
        \begin{equation*}
            |\{ j \ | \ x \in C_j \}| = |\{ j \ | \ y \in C_j \}| \ \text{for all $x, y \in L_i$}
        \end{equation*}
    \end{itemize}
\end{proposition}
\begin{proof}
    Show the directions (ii) $\Rightarrow$ (iii) $\Rightarrow$ (i) $\Rightarrow$ (ii)
    \paragraph{(ii) $\Rightarrow$ (iii)} Use induction on $n$.
    The base case is trivial, just take chains $\{x\}$ for each $x \in A = L_0$.
    So assume that $n > 0$. The set
    \begin{equation*}
        A' := \bigcup_{i = 0}^{n - 1} L_i
    \end{equation*}
    is a graded poset of maximal rank $n - 1$, and hence there are maximal chains $C_1, ..., C_t \subseteq A'$ such that for all $1 \leq i \leq n - 1$ have
    \begin{equation*}
        |\{j \ | \ x \in C_j\}| = |\{j \ | \ y \in C_j\}| \ \text{for all $x, y \in L_i$}
    \end{equation*}
    by induction hypothesis.
    Let now $V_1 := \{1\} \times [t] \times L_n$ and $V_2 := \{2\} \times [t] \times L_n$ and consider the bipartite graph $G := (V_1 \sqcup V_2, E)$ where $E$ is defined as follows:
    \begin{equation*}
        \{ \underbrace{(1, i, a)}_{\in V_1}, \underbrace{(2, j, b)}_{\in V_2} \} \in E \ \Leftrightarrow \ \max C_j < a
    \end{equation*}
    We use Hall's theorem to show that $G$ has a matching.

    Consider any $W = \{ (1, i_1, a_1), ..., (1, i_w, a_w) \} \subseteq V_1$ and let $F = \{ a_1, ..., a_w \}$.
    Then
    \begin{equation*}
        W \subseteq [t] \times F \ \Rightarrow \ |W| \leq t|F|
    \end{equation*}
    By choice of $C_1, ..., C_t$, we have that the number of $j$ with $x \in C_j$ is the same for all $x \in L_{n - 1}$, say $k$.
    Since the $C_i$ are maximal chains, each intersects $L_{n - 1}$ in exactly one element.
    So have bijection
    \begin{align*}
        \{ (x, i) \ | \ x \in L_{n - 1}, 1 \leq i \leq t, x \in C_i\} \to \{ i \ | \ 1 \leq i \leq t \}, \quad (x, i) \mapsto i
    \end{align*}
    where the set on the left-hand side has size $k |L_{n - 1}|$. It follows that $k|L_{n - 1}| = t$.

    Since $\max C_j \in L_{n - 1}$ for all $j$, we have
    \begin{align*}
        N(W) =& \{ (2, j, b) \ | \ b \in L_n, \exists a \in F: \max C_j < a \} \\
        =& \{2\} \times \{ j \ | \ \exists a \in F: \max C_j < a \} \times L_n \\
        =& \{2\} \times \{ j \ | \ \max C_j \in \partial F \} \times L_n
    \end{align*}
    and so by the above
    \begin{align*}
        |N(W)| = |L_n| \sum_{\max C_j \in \partial F} 1 = |L_n| \sum_{x \in \partial F} \sum_{x \in C_j} 1 = |L_n| \sum_{x \in \partial F} k = |L_n| |\partial F| k
    \end{align*}
    Together with the assumption that $|\partial F| |L_n| \geq |F| |L_{n - 1}|$ we see that
    \begin{equation*}
        |W| \leq t |F| = k |L_{n - 1}| |F| \leq k |\partial F| |L_n| = |N(W)|
    \end{equation*}
    So Hall's theorem yields a matching $M \subseteq E$ from $V_1$ to $V_2$.
    As $|V_1| = |V_2|$, this is already a 1-to-one correspondence.

    Now consider the sets
    \begin{equation*}
        C'_m := C_j \cup \{ a \} \ \text{for each} \ m = \{\underbrace{(1, i, a)}_{\in V_1}, \underbrace{(2, j, b)}_{\in V_2}\} \in M
    \end{equation*}
    These are chains, as $\max C_j < a$ for each $C'_m$.
    Further, for each $x, y \in L_i, i < n$ have that
    \begin{equation*}
        |\{ m \ | \ x \in C'_m \}| = |L_n \times \{ j \ | \ x \in C_j \}| = |L_n \times \{ j \ | \ y \in C_j \}| = |\{ C'_m \ | \ y \in C'_m \}|
    \end{equation*}
    as $M$ is a matching from $V_2$ to $V_1$.
    Finally, for all $x \in L_n$ we have that
    \begin{align*}
        |\{ C'_m \ | \ x \in C'_m \}| =& |\{ m \in M \ | \ \exists i, v \in V_2: m = \{ (i, x), v \} \}| = |\{ (i, x) \ | \ (i, x) \in V_1 \}| = t
    \end{align*}
    as $M$ is a matching from $V_1$ to $V_2$.

    \paragraph{(iii) $\Rightarrow$ (i)} Let $C_1, ..., C_t$ be a sequence of maximal chains given by the assumption.
    For $1 \leq i \leq n$ let $k_i$ be the number of different $j$ such that a fixed element $x \in L_i$ is contained in exactly the $C_j$.
    By assumption, this does not depend on $x$.

    As in the direction above, we see that $k_i |L_i| = t$ because each $C_j$ intersects $L_i$ in exactly one element, so there is a bijection
    \begin{align*}
        \{ (x, j) \ | \ x \in L_i, 1 \leq j \leq t, x \in C_j\} \to \{ j \ | \ 1 \leq j \leq t \}, \quad (x, j) \mapsto j
    \end{align*}
    and the set on the left has size $k_i |L_i|$.

    Since each $C_j$ is a chain and $A$ is an antichain, we find that $A$ and $C_j$ intersect in at most one element.
    So
    \begin{align*}
        t \geq& |\{ j \ | \ A \cap C_j \neq \emptyset \}| = \sum_{a \in A} |\{ j \ | \ a \in C_j \}| = \sum_{i = 0}^n \ \sum_{a \in A \cap L_i} |\{ j \ | \ a \in C_j \}| \\
        =& \sum_{i = 0}^n \ \sum_{a \in A \cap L_i} k_i = \sum_{i = 0}^n k_i |A \cap L_i| = \sum_{i = 0}^n \frac {|A \cap L_i|} {|L_i|} t
    \end{align*}
    The claim follows by canceling $t$.

    \paragraph{(i) $\Rightarrow$ (ii)}
    Consider $A := F \cup (L_{i - 1} \setminus \partial F)$.
    This is clearly an antichain, as for $x \in F, y \in L_{i - 1} \setminus \partial F$ have $y \not< x$.
    However, $r(y) < r(x)$ so also $x \not< y$, thus $x$ and $y$ are incomparable.
    Clearly elements from the same layer are incomparable.

    Thus, the assumption yields that
    \begin{equation*}
        \frac {|L_{i - 1}| - |\partial F|} {|L_{i - 1}|} + \frac {|F|} {|L_i|} = \frac {|A \cap L_{i - 1}|} {|L_{i - 1}|} + \frac {|A \cap L_i|} {|L_i|} = \sum_{j = 0}^n \frac {|A \cap L_j|} {|L_j|} \leq 1
    \end{equation*}
    This gives
    \begin{equation*}
        1 + \frac {|F|} {|L_i|} \leq 1 + \frac {|\partial F|} {|L_{i - 1}|}
    \end{equation*}
    and the claim follows.
\end{proof}

\section{Part III}
In this part, we show that sufficiently symmetric posets satisfy the conditions introduced in the previous part.
We even show a slightly stronger stronger statement than required in the task, namely that the implication to the second condition of Prop.~\ref{prop:characterization_lym} already holds ``locally'', i.e. for each layer separately.
\begin{proposition}
    \label{prop:biregular_implies_lym}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$ and let $0 < i \leq n$ such that
    \begin{itemize}
        \item Each element $a \in L_i$ covers the same number of elements in $L_{i - 1}$
        \item Each element $a \in L_{i - 1}$ is covered by the same number of elements in $L_i$
    \end{itemize}
    Then for each $F \subseteq L_i$ have
    \begin{equation*}
        \frac {|\partial F|} {|L_{i - 1}|} \geq \frac {|F|} {|L_i|}
    \end{equation*}
    In particular, if this is true for all $i$, then all the equivalent conditions from Prop.~\ref{prop:characterization_lym} follow.
\end{proposition}
\begin{proof}
    Assume that for all $a \in L_i$
    \begin{equation*}
        |\{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \}| = k \in \N
    \end{equation*}
    and for all $b \in L_{i - 1}$
    \begin{equation*}
        |\{ a \in L_i \ | \ \text{$a$ covers $b$} \}| = l \in \N
    \end{equation*}
    Double-counting pairs $(a, b) \in L_i \times L_{i - 1}$ such that $a$ covers $b$ yields $k|L_i| = l|L_{i - 1}|$.
    Now consider $F \subseteq L_i$.
    Note that by definition of $\partial F$, we have for all $a \in F$ that
    \begin{equation*}
        \{ b \in \partial F \ | \ \text{$a$ covers $b$} \} = \{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \} = \partial \{a\} \subseteq \partial F
    \end{equation*}
    Again by double-counting pairs $(a, b) \in F \times \partial F$ such that $a$ covers $b$, we find
    \begin{align*}
        k|F| =& \sum_{a \in F} k = \sum_{a \in F} |\{ b \in L_{i - 1} \ | \ \text{$a$ covers $b$} \}| = \sum_{a \in F} |\{ b \in \partial F \ | \ \text{$a$ covers $b$} \}| \\
        =& |\{ (a, b) \in F \times \partial F \ | \ \text{$a$ covers $b$} \}| = \sum_{b \in \partial F} |\{ a \in F \ | \ \text{$a$ covers $b$}\}| \\
        \leq& \sum_{b \in \partial F} |\{ a \in L_i \ | \ \text{$a$ covers $b$}\}| = \sum_{b \in \partial F} l = l |\partial F|
    \end{align*}
    Hence
    \begin{equation*}
        \frac {|\partial F|} {|L_{i - 1}|} = \frac {l |\partial F|} {l |L_{i - 1}|} \leq \frac {k |F|} {k |L_i|} = \frac {|F|} {|L_i|}
    \end{equation*}
\end{proof}

\section{Part IV}
Now we will present the first example of this Miniproject, namely the collection of partitions of $[m]$, ordered by refinement.
\begin{example}
    Let $\Pi_m = \{ \pi \subseteq \powerset(m) \ | \ \text{$\pi$ partition} \}$ be the poset of partitions ordered by refinement.
    Then $\Pi_m$ is a graded poset with rank function
    \begin{equation*}
        r: \Pi_m \to \{ 0, ..., m - 1 \}, \quad \pi \mapsto m - |\pi|
    \end{equation*}
\end{example}
\begin{proof}
    First of all, have $1 \leq |X| \leq n$ for all subsets $X \subseteq \powerset(n)$, hence the function $r$ is well-defined.
    Next we show that for all $x < y$ have $r(x) < r(y)$.
    However, if $x$ is a proper refinement of $y$, then clearly $|x| > |y|$, so $r(x) < r(y)$.

    Now assume there is a maximal chain $\pi_1 < ... < \pi_t$ in $\Pi_m$, and want to show that $t = m$.
    As the chain is maximal and $\Pi_m$ has the largest element $[m]$, we may assume that $\pi_t = [m]$ and thus $r(\pi_t) = m - 1$.
    Similarly, we may assume that $p_1 = \{ \{a\} \ | \ a \in [m]\}$ and so $r(\pi_1) = 0$.
    So it suffices to show that $r(\pi_i) = 1 + r(\pi_{i - 1})$ for all $0 < i \leq t$.

    Assume not, i.e. $|\pi_{i - 1}| \geq 2 + |\pi_i|$ for some $0 < i \leq t$.
    Since $\pi_{i - 1}$ is a refinement of $\pi_i$, we find that
    \begin{equation*}
        \pi_{i - 1} = \{ A_{11}, ..., A_{1m_1}, \ ..., \ A_{n1}, ..., A_{nm_n} \} \quad \text{with} \quad \pi_i = \Bigl\{ \bigcup_{j \leq m_1} A_{1j}, ..., \bigcup_{j \leq m_n} A_{nj} \Bigr\}
    \end{equation*}
    where $n = |\pi_i|$ and $\sum_j m_j = |\pi_{i - 1}| \geq 2 + n$.
    Hence, there is one $s \leq n$ with $m_s \geq 3$ or two different $s, r \leq n$ with $m_s, m_r \geq 2$.

    In the first case, consider
    \begin{equation*}
        \tilde{\pi} := \{ A_{jk} \ | \ j \neq s \} \cup \Bigl\{ A_{s1}, \bigcup_{2 \leq j \leq m_s} A_{sj} \Bigr\}
    \end{equation*}
    and in the second case, consider
    \begin{equation*}
        \tilde{\pi} := \{ A_{jk} \ | \ j \neq s \} \cup \Bigl\{ \bigcup_{j \leq m_s} A_{sj} \Bigr\}
    \end{equation*}
    Then $\pi_{i - 1} < \tilde{\pi} < \pi_i$ and we found a longer chain
    \begin{equation*}
        \pi_1 < ... < \pi_{i - 1} < \tilde{\pi} < \pi_i < ... < \pi_t
    \end{equation*}
    which contradicts the assumed maximality. The claim follows.
\end{proof}
\begin{proposition}
    \label{prop:partitions_recursion}
    For the number $S(m, k) := |\{ \pi \in \Pi_m \ | \ |\pi| = k\}|$ we have recursion
    \begin{equation*}
        S(m, 1) = 1 \quad \text{and} \quad S(m, k) = \sum_{n = k - 1}^{m - 1} {m - 1 \choose n} S(n, k - 1)
    \end{equation*}
    where $1 \leq k \leq m$.
\end{proposition}
\begin{proof}
    Use induction on $k$. 
    The base case $k = 1$ is trivial, so let $k > 1$.
    For any partition $\pi \in \Pi_m$, denote by $A(1, \pi)$ the unique set $A \in \pi$ with $1 \in A$.
    We find
    \begin{align*}
        S(m, k) =& |\{ \pi \in \Pi_m \ | \ |\pi| = k \}| = \Bigl| \bigcup_{\emptyset \neq A \subseteq [m], \ 1 \in A} \{ \pi \in \Pi_m \ | \ |\pi| = k, A(1, \pi) = A \} \Bigr| \\
        =& \sum_{\emptyset \neq A \subseteq [m], \ 1 \in A} |\{ \pi \in \Pi_m \ | \ |\pi| = k, A(1, \pi) = A \}| \\
        =& \sum_{\emptyset \neq A \subseteq [m], \ 1 \in A} |\{ \pi \in \Pi_m \ | \ |\pi \setminus \{A\}| = k - 1, A(1, \pi) = A\}| \\
        =& \sum_{\emptyset \neq A \subseteq [m], \ 1 \in A} |\{ \pi \in \Pi_{m - |A|} \ | \ |\pi| = k - 1 \}| \\
        =& \sum_{i = 1}^m \ \sum_{A \subseteq [m], \ |A| = i, \ 1 \in A} |\{ \pi \in \Pi_{m - i} \ | \ |\pi| = k - 1 \} \\
        =& \sum_{i = 1}^m {m - 1 \choose i - 1} S(m - i, k - 1) = \sum_{i = 0}^{m - 1} {m - 1 \choose m - i - 1} S(i, k - 1) \\
        =& \sum_{i = 1}^{m - 1} {m - 1 \choose i} S(i, k - 1) = \sum_{n = k - 1}^{m - 1} {m - 1 \choose n} S(n, k - 1)
    \end{align*}
    as for $k > 1$ have that $S(0, k) = 0$ and $S(m, k) = 0$ if $m < k$.
\end{proof}
\begin{example}
    \label{ex:partitions_have_no_lym}
    Have for $m \geq 3$ that
    \begin{equation*}
        S(m, 2) = 2^{m - 1} - 1 \quad \text{and} \quad S(m, 3) = \frac 1 2 (3^{m - 1} - 2^m + 1)
    \end{equation*}

    Note that this is the partition into \emph{exactly} $k$ sets; 
    I am not completely sure what the exercise description tells us to show, but if it is about the number of ways of decomposing $[m]$ into \emph{at most} $k$ sets, then have
    \begin{equation*}
        \Sigma S(m, k)(2) = 2^{m - 1} \quad \text{and} \quad \Sigma S(m, k)(3) = 3^{m - 1} = \frac 1 2 3^{m - 1} + \frac 1 2
    \end{equation*}
    for $m \geq 3$.
    Here $\Sigma S(m, k) = \sum_{l \leq k} S(m, l)$ is the number of partitions of $[m]$ into at most $k$ sets.
    \begin{proof}
        By Prop.~\ref{prop:partitions_recursion} have that for $m \geq 1$ we have $S(m, 1) = 1$.
        Further, for $m \geq 2$ have
        \begin{equation*}
            S(m, 2) = \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 = 2^{m - 1} - 1
        \end{equation*}
        Applying this once more yields for $m \geq 3$ that
        \begin{align*}
            S(m, 3) =& \sum_{n = 1}^{m - 1} {m - 1 \choose n} (2^{n - 1} - 1) = \frac 1 2 \sum_{n = 1}^{m - 1} {m - 1 \choose n} 2^n - \sum_{n = 1}^{m - 1} {m - 1 \choose n} 1 \\
            =&\frac 1 2 \sum_{n = 0}^{m - 1} {m - 1 \choose n} 2^n - \frac 1 2 - \sum_{n = 0}^{m - 1} {m - 1 \choose n} 1 + 1 \\
            =& \frac 1 2 (1 + 2)^{m - 1} - 2^{m - 1} + \frac 1 2 = \frac 1 2 (3^{m - 1} - 2^m + 1)
        \end{align*}
    \end{proof}
    Let $m$ be even and consider $\mathcal{A} \subseteq \Pi_m$ of partitions of $[m]$ into two equally sized sets.
    Then
    \begin{equation*}
        |\mathcal{A}| = \frac 1 2 {m \choose m/2}
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{equation*}
            f: [m]^{(m/2)} \to \mathcal{A}, \quad A \mapsto \{ A, [m] \setminus A \}
        \end{equation*}
        Then for $A$ and $B \neq A, [m] \setminus A$ have that $\{ A, [m] \setminus A \} \neq \{ B, [m] \setminus B \}$.
        Conversely, $A$ and $[m] \setminus A$ have the same image under $f$.
        This shows that $f$ is 2-to-1 and thus 
        \begin{equation*}
            |\mathcal{A}| = \frac 1 2 \left| [m]^{(m/2)} \right| = \frac 1 2 {m \choose m/2}
        \end{equation*}
    \end{proof}
    Further have
    \begin{equation*}
        |\partial \mathcal{A}| = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
    \end{equation*}
    \begin{proof}
        Consider the map
        \begin{align*}
            g: \underbrace{\{ (A, B, C) \ | \ A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m \}}_{ =: \mathcal{G} } \to \partial \mathcal{A}, \quad (A, B, C) &\mapsto \{ A, B, C \}
        \end{align*}
        Note that $g$ is well-defined, as for $A \in [m]^{(m/2)}, \{ A, B, C \} \in \Pi_m$ have that $\pi := \{ A, B, C \}$ is a partition in $L_{m - 3}$ that refines the partition $\{ A, B \cup C \} \in \mathcal{A} \subseteq L_{m - 2}$.
        Further, it is easy to see that $g$ is surjective, as every partition $\pi \in \partial \mathcal{A}$ satisfies $|\pi| = 3$ and has some $A \in \pi$ with $|A| = m/2$.

        To complete the proof, we investigate to what ``extend $g$ is injective''.

        Assume $g(A, B, C) = g(A', B', C')$. As $B, C, B', C' \neq \emptyset$ and $|B \cup C| = |B' \cup C'| = m/2$, we see that $A = A'$.
        Further, we must then have that $\{ B, C \} = \{ B', C' \}$, hence $(A, B, C) = (A', B', C')$ or $(A, B, C) = (A', C', B')$ (clearly $B \neq C, B' \neq C'$).
        This shows that the map $g$ is 2-to-1.

        Using this, we find
        \begin{align*}
            |\partial \mathcal{A}| =& \frac 1 2 |\mathcal{G}| = \frac 1 2 \sum_{A \in [m]^{(m/2)}} |\{ (B, C) \ | \ \{ B, C \} \ \text{partition of} \ [m] \setminus A \}| \\
            =& \frac 1 2 \sum_{A \in [m]^{(m/2)}} 2 S(m/2, 2) = \sum_{A \in [m]^{(m/2)}} 2^{m/2 - 1} - 1 = {m \choose m/2} \left( 2^{m/2 - 1} - 1 \right)
        \end{align*}
    \end{proof}
    We can now plug this into the second condition of Prop.~\ref{prop:characterization_lym} to see
    \begin{align*}
        \frac {|\partial \mathcal{A}|} {|L_{m - 3}|} &= \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {S(m, 3)} = \frac {{m \choose m/2}\left( 2^{m/2 - 1} - 1\right)} {\frac 1 2 \left( 3^{m - 1} - 2^m + 1 \right)} \\
        &\quad \sim \frac {2^{m + m/2}} {\sqrt{m}(3^{m - 1} - 2^m)} \sim \frac {3 \left( 2^{3/2} \right)^m} {\sqrt{m} 3^m} = \frac 3 {\sqrt{m}} c^m
    \end{align*}
    and
    \begin{align*}
        \frac {|\mathcal{A}|} {|L_{m - 2}|} &= \frac {\frac 1 2 {m \choose m/2}} {S(m, 2)} = \frac {{m \choose m/2}} {2^m - 2} \sim \frac {2^m} {\sqrt{m}2^m} = \frac 1 {\sqrt{m}}
    \end{align*}
    for some $0 < c < 1$ (here $\sim$ means asymptotically equivalent as $m \to \infty$).
    In particular, find for all $\epsilon > 0$ and sufficiently large $m$ that
    \begin{equation*}
        \frac {|\partial \mathcal{A}|} {|L_{m - 3}|} < \epsilon \frac {|\mathcal{A}|} {|L_{m - 2}|}
    \end{equation*}
    and so the conditions of Prop.~\ref{prop:characterization_lym} are not satisfied.
\end{example}

\section{Part V}
Now we come to the second example, namely multi-subsets of a $d$-element multiset, ordered by containment.
Note that we can characterize multisets in different ways, e.g. as vectors $\N^d$ ordered elementwise, or as natural numbers ordered by divisibility.
In this section, we will focus on the former perspective, but in the last part, we will have a short remark on the latter view.
\begin{example}
    Let $\mathcal{P}_{k, d} := \{ 0, ..., k \}^d$ partially ordered by elementwise ordering.
    Then $\mathcal{P}_{k, d}$ is a graded poset with rank function
    \begin{equation*}
        r: \mathcal{P}_{k, d} \to \{ 0, ..., kd \}, \quad a \mapsto \sum_i a_i
    \end{equation*}
\end{example}
\begin{proof}
    First, we show that for $a < b$ have $r(a) < r(b)$.
    If $a < b$, then we find that $a_j \leq b_j$ for all $j$ and $a_i \neq b_i$, so $a_i < b_i$ for some $i$.
    Hence
    \begin{equation*}
        r(a) = \sum_j a_j = a_i + \sum_{j \neq i} a_j \leq a_i + \sum_{j \neq i} b_j < b_i + \sum_{j \neq i} b_j = \sum_j b_j = r(b)
    \end{equation*}
    Now assume that we have a maximal chain $a_1 < ... < a_t$ in $\mathcal{P}_{k, d}$.
    Since $\mathcal{P}_{k, d}$ has the smallest element $0 = (0, ..., 0)$ and the largest element $k = (k, ..., k)$, we see that $p_1 = 0$ and $p_t = k$.
    We want to show $t = kd + 1$, so it suffices to show that $r(a_i) = r(a_{i - 1}) + 1$ for all $0 < i \leq t$.

    Assume not, then 
    \begin{equation*}
        r(a_i) - r(a_{i - 1}) = \sum_j a_{ij} - a_{(i - 1)j} \geq 2
    \end{equation*}
    For all $j$ we have $a_{ij} \geq a_{(i - 1)j}$ and so there is $s \leq d$ with $a_{is} - a_{(i - 1)s} \geq 2$ or there are different $r, s \leq d$ with $a_{ir} - a_{(i - 1)r}, a_{is} - a_{(i - 1)s} \geq 1$.
    In both cases, find
    \begin{equation*}
        \tilde{a} \in \mathcal{P}_{k, d} \ \text{defined by} \ \tilde{a}_j = \begin{cases}
            a_{(i - 1)j} & \text{if $j \neq s$} \\
            a_{(i - 1)j} + 1 & \text{otherwise}
        \end{cases}
    \end{equation*}
    with $a_{i - 1} < \tilde{a} < a_i$. However, this gives a longer chain
    \begin{equation*}
        a_1 < ... < a_{i - 1} < \tilde{a} < a_i < ... < a_t
    \end{equation*}
    which contradicts the assumed maximality. The claim follows.
\end{proof}
\begin{definition}
    Let $P$ be a graded poset with layers $L_0, ..., L_n$.
    A symmetric chain in $P$ is a chain $C \subseteq P$ such that there is $i \leq n/2$ with
    \begin{equation*}
        \forall 0 \leq j \leq n: \ C \cap L_j \neq \emptyset \ \Leftrightarrow \ i \leq j \leq n - i
    \end{equation*}
    A decomposition $\mathcal{C}$ of $P$ into chains is called symmetric chain decomposition, if all $C \in \mathcal{C}$ are symmetric.
\end{definition}
Note that this definition is compatible with the definition of a symmetric chain for the classical graded poset $\powerset(n)$.
\begin{proposition}
    Let $k, d \geq 1$. Then $\mathcal{P}_{k, d}$ has a symmetric chain decomposition.
\end{proposition}
\begin{proof}
    Use induction on $d$.
    The base case $d = 1$ is trivial, so assume that $d > 1$.
    By induction hypothesis, have a symmetric chain decomposition $\mathcal{C}$ of $\mathcal{P}_{k, d - 1}$.
    Now consider some $C = \{ A_i, A_{i + 1}, ..., A_{d(k - 1) - i} \} \in \mathcal{C}, i < \frac n 2$ with $r(A_j) = j$.
    For each $0 \leq l \leq k, n - 2i - 1$ define the chain
    \begin{equation*}
        \tilde{C}_l := \{ \underbrace{(A_{i + l}, j)}_{\mathclap{:= (a_1, ..., a_{d - 1}, j) \ \text{if} \ A_{i + l} = (a_1, ..., a_{d - 1})}} \ | \ 0 \leq j \leq k - l \} \cup \{ (A_j, k - l) \ | \ i + l < j \leq d(k - 1) - (i + l) \}
    \end{equation*}
    Then $\tilde{C}_l$ is symmetric w.r.t. $i + l$, as for $i + l \leq j \leq dk - (i + l)$ we have
    \begin{align*}
        i + l \leq j \leq k + i \quad &\Rightarrow \quad (A_{i + l}, j - i - l) \in \tilde{C}_l, \\
        k + i < j \leq dk - (i + l) \quad &\Rightarrow \quad (A_{j - k}, k - l) \in \tilde{C}_l
    \end{align*}
    Clearly the $\tilde{C}_l$ are chains and disjoint.
    Hence, we find a symmetric chain decomposition
    \begin{equation*}
        \tilde{\mathcal{C}} := \Bigl\{ \tilde{C}_l \ \Bigm| \ C \in \mathcal{C} \ \text{symmetric chain w.r.t. $i$}, \ 0 \leq l \leq \min\{ k, n - 2i - 1 \} \Bigr\}
    \end{equation*}
    of $\mathcal{P}_{k, d}$.
\end{proof}
Note that for $k = 1$, this is exactly the proof that was done in the lecture to show that $\powerset(d)$ has a symmetric chain decomposition.

\section{Part VI}
First of all, in Example~\ref{ex:partitions_have_no_lym} we have already shown that the poset $\Pi_m$ does not satisfy the conditions from Prop.~\ref{prop:characterization_lym}.
Now we want to investigate whether $\mathcal{P}_{k, d}$ does. 
\begin{lemma}
    \label{prop:lp_algorithm}
    Let $P$ be a poset with layers $L_0, ..., L_d$.
    Fix $0 < l \leq d$ and let $L_{l - 1} = \{ u_1, ..., u_m \}, L_l = \{ v_1, ..., v_n \}$.

    Assume there exists a matrix $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ such that
    \footnote{We use $1 = (1, ..., 1)^T$ as the vector consisting of $1$ only here.}
    \begin{equation*}
        1^T A 1 = 1
    \end{equation*}
    and for all $i, j$
    \begin{equation*}
        e_i^T A 1 = \frac 1 {|L_0|} \quad \text{and} \quad 1^T A e_j = \frac 1 {|L_1|}
    \end{equation*}
    and $a_{ij} = 0$ whenever $u_i \not< v_j$ holds.

    Then the conditions of Prop.~\ref{prop:characterization_lym} hold for the poset $L_{l - 1} \cup L_l$.
    In particular, if the assumption holds for all $0 < l \leq d$, then the conditions of Prop.~\ref{prop:characterization_lym} hold for $P$.
\end{lemma}
\begin{proof}
    Let $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ be this matrix.
    Note that a maximal chain in $L_{l - 1} \cup L_l$ corresponds to a pair $u_i < v_j$.

    Now consider a (finite) sequence of maximal chains $(C_k)_{k \in I}$ such that the chain corresponding to $u_i < v_j$ occurs $a_{ij} |I|$ times.
    By the assumption, we know that for each $i$ have
    \begin{equation*}
        |\{k \ | \ u_i \in C_k\}| = \sum_{u_i < v_j} a_{ij}|I| = \frac 1 {|L_0|} |I|
    \end{equation*}
    and so each element $u_i \in L_{l - 1}$ is contained in the same number $|I| / |L_0|$ of chains $C_k$.

    Similarly, for each $J$ have
    \begin{equation*}
        |\{k \ | \ v_j \in C_k\}| = \sum_{v_j > u_i} a_{ij}|I| = \frac 1 {|L_1|} |I|
    \end{equation*}
    The claim follows.
\end{proof}
This lemma does not seem super useful at first, but it gives us a way to check in time polynomial in $O(|P|)$ whether a given poset $P$ satisfies the conditions from Prop.~\ref{prop:characterization_lym}, by using an LP solver.
Note that this is not easily possible when working directly with any of the three conditions defined before.

Using this, I could check that $\mathcal{P}_{k, d}$ indeed satisfies the conditions for reasonably sized $|\mathcal{P}_{k, d}|$ (up to $k = d = 6$).
Note that already for these ``small'' numbers, we get quite huge posets $|\mathcal{P}_{k, d}|$.
Hence, I believe that this is true for all $\mathcal{P}_{k, d}$.

My successful-looking attempt at proving this is based on the observation that elements of $\mathcal{P}_{k, d}$ ``behave similarly'' (in the sense that there is an element of the automorphism group $\mathrm{Aut}(\mathcal{P}_{k, d})$ mapping one to the other) if the have the same number of equally sized entries.
This looks promising, as the proof for $\powerset(n)$ relies on the fact that all elements in one layer of $\powerset(n)$ ``behave similarly'' in this sense.
In particular, note that Prop.~\ref{prop:biregular_implies_lym} would already be applicable in such a case.

This is formalized in the next definition.
\begin{definition}
    Let $P = \mathcal{P}_{k, d}$. Then define
    \begin{equation*}
        P(n_0, ..., n_k) := \bigl\{ x \in P \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j \bigl\}
    \end{equation*}
    Further, for any subset $F \subseteq P$ write
    \begin{equation*}
        F(n_0, ..., n_k) := F \cap P(n_0, ..., n_k)
    \end{equation*}
\end{definition}
\begin{lemma}
    \label{prop:properties_n_distribution_subsets}
    The size of $|P(n_0, ..., n_k)|$ is given by the multinomial coefficient
    \begin{equation*}
        |P(n_0, ..., n_k)| = { d \choose n_0, ..., n_k } =  \frac {d!} {n_0! ... n_k!}
    \end{equation*}
    if $n_0 + ... + n_k = d$ and $n_i \geq 0$ (otherwise, it is clearly zero).

    Further, for all $n_0, ..., n_k \in \Z$ with $\sum n_i = d$ and all $a \in P(n_0, ..., n_k), 0 \leq \delta < k$ have
    \begin{align*}
        |\partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)| &= n_{\delta + 1}, \\
        |\partial^+ a \cap P(n_0, ..., n_\delta - 1, n_{\delta + 1} + 1, ..., n_k)| &= n_\delta
    \end{align*}
    even if any of the $n_i$ are negative.
\end{lemma}
\begin{proof}
    For $n_0 + ... + n_k = d$ and $n_i \geq 0$ we have
    \begin{align*}
        &|P(n_0, ..., n_k)| = |\bigl\{ x \in P \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j\bigl\}| \\
        =& \sum_{A_0 \in [d]^{(n_0)}} |\bigl\{ x \in \{0, ..., k\}^{[d] \setminus A_0} \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j\bigl\}| \\
        =& \sum_{A_0 \in [d]^{(n_0)}} \sum_{A_1 \in ([d] \setminus A_0)^{(n_1)}} |\bigl\{ x \in \{0, ..., k\}^{[d] \setminus A_0 \setminus A_1} \ \bigm| \ |\{i \ | \ x(i) = j\}| = n_j\bigl\}| \\
        =& \ ... \\
        =& \sum_{A_0 \in [d]^{(n_0)}} ... \sum_{A_k \in ([d] \setminus A_0 \setminus ... \setminus A_{k - 1})^{(n_k)}} 1 \\
        =& \sum_{A_0 \in [d]^{(n_0)}} ... \sum_{A_{k - 1} \in ([d] \setminus A_0 \setminus ... \setminus A_{k - 2})^{(n_k)}} {d - n_0 - ... - n_{k - 1} \choose n_k} \\
        =& \ ... \\
        =& {d \choose n_0} {d - n_0 \choose n_1} ... {d - n_0 - ... - n_{k - 1} \choose n_k} = \prod_i \frac {(d - n_0 - ... - n_{i - 1})!} {n_i! (d - n_0 - ... - n_i!)} \\
        =& \frac {d!} {n_0! ... n_k!} \prod_i \frac {(d - n_0 - ... - n_i)!} {(d - n_0 - ... - n_i)!} = \frac {d!} {n_0! ... n_k!}
    \end{align*}
    Further have for all $n_0 + ... + n_k = d$ and $a \in P(n_0, ..., n_k)$ that there is a bijection
    \begin{align*}
        \{ i \in [d] \ | \ a(i) = \delta + 1 \} &\to \partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k), \\
        i &\mapsto f_{a, i}: j \mapsto \begin{cases}
            a(j) & \text{if $j \neq i$} \\
            a(j) - 1 & \text{otherwise}
        \end{cases}
    \end{align*}
    It is easy to see that the image of some $i$ is indeed an element of $\partial a$ and of $P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)$.
    Clearly the function is also injective, and surjectivity can be shown by considering some $b \in \partial a \cap P(n_0, ..., n_\delta + 1, n_{\delta + 1} - 1, ..., n_k)$ and showing by induction that
    \begin{equation*}
        \{ i \ | \ b(i) = j \} = \{ i \ | \ a(i) = j \}
    \end{equation*}
    for all $j \neq \delta, \delta + 1$.
    Note that the set $\{ i \in [d] \ | \ a(i) = \delta + 1 \}$ has size exactly $n_{\delta + 1}$ and so the claim follows.

    The analogue statement for the upper shadow of $a$ can be proven in the same way.
\end{proof}
\begin{definition}
    Define the \emph{core} of $\mathcal{P}_{k, d}$ as the poset
    \begin{equation*}
        \mathcal{C}_{k, d} := \{ (n_0, ..., n_k) \ | \ n_0 + ... + n_k = d \}
    \end{equation*}
    whose partial order is the transitive closure of the relation $\prec$ defined by
    \begin{align*}
        (n_0, ..., n_k) \prec (m_0, ..., m_k) \ :\Leftrightarrow \ \exists 0 \leq \delta < k: \ &m_\delta = n_\delta - 1, \ m_{\delta + 1} = n_{\delta + 1} - 1, \\
        &\forall i \neq \delta, \delta + 1: m_i = n_i 
    \end{align*}
\end{definition}
\begin{lemma}
    The layers of $\mathcal{C}_{k, d}$ are
    \begin{equation*}
        L_i = \{ (n_0, ..., n_k) \ | \ \sum_j j n_j = i \}
    \end{equation*}
    Further, we have a surjective map
    \begin{equation*}
        R: \mathcal{P}_{k, d} \to \mathcal{C}_{k, d}, \quad A \mapsto \bigl( |\{j \ | \ A(j) = i\}| \bigr)_{0 \leq i \leq k}
    \end{equation*}
    that preserves the ordering and the fibre of a point $(n_0, ..., n_k)$ is exactly $P(n_0, ..., n_k)$.
\end{lemma}
\begin{proof}
    All of these statements are very easy to check.
\end{proof}

\begin{proposition}
    \label{prop:core_of_multisets}
    Assume that $\mathcal{C}_{k, d}$ satisfies a ``scaled variant'' of condition (3) of Prop.~\ref{prop:characterization_lym}, namely:
    There exists a (finite) sequence $C'_1, ..., C'_t$ of maximal chains such that for a chain $C'$ chosen uniformly at random among them, we have
    \begin{equation*}
        \Pr[ (n_0, ..., n_k) \in C' ] = \frac {|P(n_0, ..., n_k)|} {|L_{r(n_0, ..., n_k)}|}
    \end{equation*}
    Then $\mathcal{P}_{k, d}$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}.
\end{proposition}
\begin{proof}
    Denote the layers of $\mathcal{P}_{k, d}$ by $L_k$ and the layers of $\mathcal{C}_{k, d}$ by $L'_k$.

    Now consider a chain $C'_i$ from the assumption.
    Note that by Lemma~\ref{prop:properties_n_distribution_subsets}, we see that for $(n_0, ..., n_k) \in C'_i$ have
    \begin{itemize}
        \item Every element in $P(n_0, ..., n_k)$ is covered by $n_\delta$ elements in $P'_i$
        \item Every element in $P(n_0, ..., n_k)$ covers $n_{\delta + 1}$ elements in $P'_i$
    \end{itemize}
    where $P'_i \subseteq P$ is the graded poset that is given by the preimage of $C'_i$ under the canonical map $\mathcal{P}_{k, d} \to \mathcal{C}_{k, d}$, i.e.
    \begin{equation*}
        P'_i := \bigcup_{c \in C'_i} P(c) = R^{-1}(C'_i)
    \end{equation*}
    The $P(n_0, ..., n_k)$ are exactly the layers of the poset $P'_i$.
    Hence Prop.~\ref{prop:biregular_implies_lym} yields that there is a sequence of maximal chains
    \begin{equation*}
        C_{i, 1}, ..., C_{i, t_i}
    \end{equation*}
    in the poset $P'_i$ such that each element in $P(n_0, ..., n_k)$ is contained in $t_i/|P(n_0, ..., n_k)|$ of them.
    Note that $R(C_{i, j}) := \{ R(x) \ | \ x \in C_{i, j} \} = C'_i$ holds.

    Now consider a sequence $(C_{i_h, j_h})_{h \in I}$ of maximal chains in the poset $A'$ that consists of the $C_{i, j}$ such that one $C_{i, j}$ occurs $\frac 1 {t t_i} |I|$ times.
    Let $H$ be a random variable distributed uniformly on $I$.
    Observe that then $i_H$ is a random variable distributed uniformly on $1, ..., t$.

    Let $x \in L_k$ with $x \in P(n_0, ..., n_k)$.
    Then
    \begin{align*}
        \Pr[x \in C_{i_H, j_H}] =& \sum_{1 \leq i \leq t} \Pr[x \in C_{i_H, j_H} \ | \ i_H = i] \ \Pr[i_H = i] \\
        =& \sum_{1 \leq i \leq t} \underbrace{\Pr[x \in C_{i_H, j_H} \ | \ i_H = i]}_{ = 1 / |P(n_0, ..., n_k)| \ \text{if $(n_0, ..., n_k) \in C'_i$}} \frac 1 t \\
        =& \sum_{\substack{1 \leq i \leq t\\(n_0, ..., n_k) \in C'_i}} \frac 1 {|P(n_0, ..., n_k)| t} \\
        =& \frac t {|P(n_0, ..., n_k)| t} \underbrace{\Pr[(n_0, ..., n_k) \in C'_{i_H}]}_{= |P(n_0, ..., n_k)| / |L_k| \ \text{by assumption}} = \frac 1 {|L_k|}
    \end{align*}

    Clearly, this shows that condition (3) of Prop.~\ref{prop:characterization_lym} is satisfied by $(C_{i_h, j_h})_{h \in I}$ for $\mathcal{P}_{k, d}$.
\end{proof}
After some reflection, one can observe that we did not really use any properties of $\mathcal{P}_{k, d}$.
The fact that all elements in $P(n_0, ..., n_k)$ cover resp. are covered by the same number of elements requires only that $P(n_0, ..., n_k)$ is an orbit under the action of $\mathrm{Aut}(P)$, i.e. $P(n_0, ..., n_k) = \mathrm{Aut}(P).a$ for any $a \in P(n_0, ..., n_k)$
(where the dot $.$ denotes the group action of $\mathrm{Aut}(P)$ on $P$).

Hence, we get the following statement.
\begin{proposition}
    Let $P$ be a graded poset. Define the \emph{core} $\mathcal{C}(P)$ of $P$ as the set of orbits under the action of $\mathrm{Aut}(P)$ 
    \begin{equation*}
        \mathcal{C}(P) := \{ \mathrm{Aut}(P).x \ | \ x \in P \}
    \end{equation*}
    with the induced partial ordering from the elements of each orbit, i.e.
    \begin{equation*}
        A < B \ :\Leftrightarrow \ \exists a \in A, b \in B: a < b
    \end{equation*}
    for $A, B \in \mathcal{C}(P)$.
    Note that then $\mathcal{C}(P)$ is a graded poset.

    If there exists a sequence of maximal chains $C'_1, ..., C'_t$ in $\mathcal{C}(P)$ such that for a chain $C$ chosen uniformly at random from among them, have
    \begin{equation*}
        \Pr[A \in C] = \frac {|A|} {|L_{r(A)}|}
    \end{equation*}
    for all $A \in \mathcal{C}(P)$, then $P$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}.
\end{proposition}
\begin{proof}
    Exactly as in Prop.~\ref{prop:core_of_multisets}.
\end{proof}
I still do not know how to use this to prove that $\mathcal{P}_{k, d}$ does indeed satisfy Prop.~\ref{prop:characterization_lym}, but at least it provides a way for an even faster algorithm to test it for specific $k, d$.
The idea is just that we can adjust the LP algorithm from Lemma~\ref{prop:lp_algorithm} to check the scaled variant, and so we get a polynomial time algorithm in $|\mathcal{C}_{k, d}|$.
In particular, we have the following.
\begin{lemma}
    Let $L_{l - 1} = \{u_1, ..., u_m\}$ and $L_l = \{v_1, ..., v_n\}$ be two consecutive layers of $\mathcal{C}_{k, d}$.
    Assume there exists a matrix $A = (a_{ij}) \in \Q_{\geq 0}^{m \times n}$ such that 
    \begin{equation*}
        1^T A 1 = 1
    \end{equation*}
    and for all $i, j$
    \begin{equation*}
        e_i^T A 1 = \frac {|P(u_i)|} {\sum_{u \in L_{l - 1}} |P(u)|} \quad \text{and} \quad 1^T A e_j = \frac {|P(v_j)|} {\sum_{v \in L_l} |P(v)|}
    \end{equation*}
    and $a_{ij} = 0$ whenever $u_i \not< v_j$ holds.

    Then the assumption of Prop.~\ref{prop:core_of_multisets} is fulfilled for $L_{l - 1} \cup L_l$.
    In particular, if this holds for all $0 < l \leq kd$ then the assumption is fulfilled for the whole of $\mathcal{C}_{k, d}$ and it follows that $\mathcal{P}_{k, d}$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}.
\end{lemma}
\begin{proof}
    Exactly as in Lemma~\ref{prop:lp_algorithm}.
\end{proof}
Using a simple implementation of this approach, I could check that for some arbitrarily chosen values of $k, d$ up to $(k, d) = (8, 9)$, the poset $\mathcal{P}_{k, d}$ indeed satisfies the conditions of Prop.~\ref{prop:characterization_lym}.
Note that for this choice of $k, d$ we already have that $|\mathcal{P}_{8, 9}| = 134217728$ which is absolutely infeasible for more naive approaches.

After quite a long time trying to prove that this holds generally, I finally gave up and searched the literature.
There I found that $\mathcal{P}_{k, d}$ is also called divisor lattice and indeed we have the following statement.
\begin{proposition}
    $\mathcal{P}_{k, d}$ satisfies the conditions of Prop.~\ref{prop:characterization_lym}, i.e. an LYM-inequality.
\end{proposition}
\begin{proof}
    The proof is similar to \cite[Theorem 4.2.3]{anderson}.
    However, I really like the proof and so we present it here in a somewhat modified form.

    We show condition (2) of Prop.~\ref{prop:characterization_lym}, so consider some pair of consecutive layers $L_{i - 1}, L_i$.
    Let further $F \subseteq L_i$. 
    We want to show that
    \begin{equation*}
        |L_{i - 1} \times F| \leq |L_i \times \partial F|
    \end{equation*}
    Let now $P' := \{ 0, ..., 2k \}^d$ and note that we have the decompositions
    \begin{align*}
        L_{i - 1} \times F &= \bigcup_{M \in P', r(M) = 2i - 1} \{ (u, a) \in L_{i - 1} \times F \ | \ u + a = M \} \quad \text{and} \\
        L_i \times \partial F &= \bigcup_{M \in P', r(M) = 2i - 1} \{ (v, b) \in L_i \times \partial F \ | \ v + b = M \}
    \end{align*}
    Hence it suffices to show that
    \begin{equation*}
        |\{ (u, a) \in L_{i - 1} \times F \ | \ u + a = M \}| \leq |\{ (v, b) \in L_i \times \partial F \ | \ v + b = M \}|
    \end{equation*}
    for each $M \in P', r(M) = 2i - 1$.

    To do so, we show the following, stronger claim by induction.
\end{proof}
\begin{lemma}
    Let $k_1, ..., k_d > 0$ and consider the poset
    \begin{equation*}
        P = \bigtimes_{i = 1}^d \{ 0, ..., k_d \}
    \end{equation*}
    Let further $M = (M_1, ..., M_d)$ with $M_i \leq 2k_i$ of rank $\sum_i M_i = 2l - 1$ and $F \subseteq L_l$ such that $M$ is an upper bound for $F$.
    Then
    \begin{equation*}
        |\{ (u, a) \in L_{l - 1} \times F \ | \ u + a = M \}| \leq |\{ (v, b) \in L_l \times \partial F \ | \ v + b = M \}|
    \end{equation*}
\end{lemma}
\begin{proof}
    Use induction on $\sum_i k_i$. 
    The base case is trivial, so assume $\sum_i k_i > 1$.

    \paragraph{Case 1} If there exists $j$ such that $M_j < k_j$ then clearly
    \begin{equation*}
        F, \partial F \in P' := \{ 0, ..., k_1 \} \times ... \times \{ 0, ..., k_j - 1 \} \times ... \times \{ 0, ..., k_d \}
    \end{equation*}
    and also $u \in P'$ for all $u \in L_{l - 1}$ with $u + a = M$ for some $a \in F$.
    Hence the induction hypothesis yields
    \begin{equation*}
        |\{ (u, a) \in L_{l - 1} \times F \ | \ u + a = M \}| \leq |\{ (v, b) \in L_l \times \partial F \ | \ v + b = M \}|
    \end{equation*}

    \paragraph{Case 2} If there exists $j$ such that $M_j > k_j \geq 1$, then consider
    \begin{align*}
        F' := &\{ a - e_j \ | \ a \in F, a(j) \neq 0 \} \in \\
        &P' := \{ 0, ..., k_1 \} \times ... \times \{ 0, ..., k_j - 1 \} \times ... \times \{ 0, ..., k_d \}
    \end{align*}
    Clearly $M - 2e_j \geq (k_1, ..., k_j - 1, ..., k_d) \geq a - e_j$ for all $a \in F$, and so $M - 2e_j$ is an upper bound for $F'$.
    By induction hypothesis, we now get that
    \begin{equation*}
        |\{ (u, a) \in L_{l - 2} \times F' \ | \ u + a = M - 2e_j \}| \leq |\{ (v, b) \in L_{l - 1} \times \partial F' \ | \ v + b = M - 2e_j \}|
    \end{equation*}
    However, observe that the map
    \begin{align*}
        \{ (u, a) \in L_{l - 2} \times F' \ | \ u + a = M - 2e_j \} &\to \{ (u, a) \in L_{l - 1} \times F \ | \ u + a = M \} \\
        (u, a) &\mapsto (u + e_j, a + e_j)
    \end{align*}
    is a bijection.
    The only nontrivial part is to see that it is surjective, so consider some $(u, a) \in L_{l - 1} \times F$ with $u + a = M$.
    Then $u(j), a(j) \leq k_j < M(j)$ and so $a(j) = M(j) - u(j) \geq 1$, hence $a - e_j \in F'$.
    Clearly $u - e_j \in L_{l - 2}$ and $u - e_j + a - e_j = M - 2e_j$.
    So $(u, a)$ has the preimage $(u - e_j, a - e_j)$.

    Similarly, see that the map
    \begin{align*}
        \{ (u, a) \in L_{l - 1} \times \partial F' \ | \ u + a = M - 2e_j \} &\to \{ (u, a) \in L_l \times \partial F \ | \ u + a = M \} \\
        (u, a) &\mapsto (u + e_j, a + e_j)
    \end{align*}
    is a bijection.
    Thus
    \begin{equation*}
        |\{ (u, a) \in L_{l - 1} \times F \ | \ u + a = M \}| \leq |\{ (v, b) \in L_l \times \partial F \ | \ v + b = M \}|
    \end{equation*}

    \paragraph{Case 3} If $M = (k_1, ..., k_d)$, see that $2l - 1 = \sum_i k_i$ and so the layers $L_{l - 1}, L_l$ must be the two unique middle layers.
    Now note that $|L_l| = |L_{l - 1}|$ and so it suffices to show that $|\partial F| \geq |F|$.
    However, this follows directly by the symmetric chain decomposition of the poset $P$.
    We have proven the symmetric chain decomposition only for $k_1 = ... = k_d$, but the general version can be found in \cite{anderson}.
\end{proof}
There are two parts I find interesting in this proof.
\begin{itemize}
    \item First of all, the idea to consider all $\bigtimes_i \{ 0, ..., k_i \}$ with possibly distinct $k_i$ makes it easy to use induction, as it suffices to decrease just one coordinate to use the induction hypothesis.
    \item The decomposition
    \begin{equation*}
        L_{i - 1} \times F = \bigcup_{M \in P', r(M) = 2i - 1} \{ (u, a) \in L_{i - 1} \times F \ | \ u + a = M \}
    \end{equation*}
    is just very clever, as it really transforms the set of tuples $(u, a)$ into multiple sets that are more or less $\{ a \in F \ | \ a \leq M \}$ and much easier to handle.
\end{itemize}
Interestingly, I have also found a variant of the LYM inequality (``pseudo-LYM inequality'') for the poset $\Pi_m$ that we studied earlier.
\begin{proposition}
    \label{prop:lym_variant_partitions}
    Let $\mathcal{A} \subseteq \Pi_m$ be a nonempty antichain. Then define
    \begin{equation*}
        a(n_1, ..., n_k) := |\bigl\{ A \in \mathcal{A} \ \bigm| \ |A| = k, \ |\{ S \in A \ | \ |S| = j \}| = n_j \bigr\}|
    \end{equation*}
    the number of partitions $A \in \mathcal{A}$ such that the classes in $A$ have sizes $n_1, ..., n_k$.
    Then
    \begin{equation*}
        \sum_{k \geq 1} \sum_{n_1, ..., n_k} \frac {a(n_1, ..., n_k) k!} {{m \choose n_1, ..., n_k} {m - 1 \choose k - 1}} \leq 1
    \end{equation*}
\end{proposition}
\begin{proof}
    See \cite{lym_partitions}.
\end{proof}
Note that the total number of partitions $\pi$ such that the classes in $\pi$ have sizes $n_1, ..., n_k$ is given by an expression involving the multinomial coefficient, namely
\begin{equation*}
    |\bigl\{ \pi \in \Pi_m \ \bigm| \ |\pi| = k, |\{ S \in \pi \ | \ |S| = j\}| = n_j \bigr\}| = {m \choose n_1, ..., n_k} \underbrace{ \prod_i \frac 1 {|\{ j \ | \ n_j = i\}|!} }_{=: \gamma(n_1, ..., n_k)}
\end{equation*}
Intuitively, the reason is that there are $m!$ ways to order the elements $[m]$, and after splitting such a sequence in chunks of sizes $n_1, ..., n_k$, there are $n_1! \cdot ... \cdot n_k!$ ways to order the element within the chunks.
Hence the multinomial coefficient
\begin{equation*}
    {m \choose n_1, ..., n_k}
\end{equation*}
gives the number of ways to distribute the elements of $[m]$ into ordered buckets of sizes $n_1, ..., n_k$.
A partition corresponds to unordered buckets, i.e. swapping chunks of equal size does not change the partition.
There are $1/\gamma(n_1, ..., n_k)$ ways to do so, and hence we get the expression mentioned above.

Furthermore, the number $a(n_1, ..., n_k)$ and the underlying set do not depend on the order of $n_1, ..., n_k$.
Hence, the partitions $A \in \mathcal{A}$ such that the classes of $A$ have sizes $n_1, ..., n_k$ might be counted multiple times.
Note that the number of different orderings of $n_1, ..., n_k$ is given by $k!/\gamma(n_1, ..., n_k)$.
Hence
\begin{equation*}
    \sum_{n_1, ..., n_k} \frac 1 {k!} {m \choose n_1, ..., n_k} = |L_{m - k}|
\end{equation*}
just counts all the partitions in one layer of $\Pi_m$.

Now note also that
\begin{equation*}
    {m - 1 \choose k - 1}
\end{equation*}
is exactly the number of different sequences $n_1 \leq ... \leq n_k$ summing to $m$ by the ``stars and bars'' method.
Then one can see that the expression
\begin{equation*}    
    \sum_{k \geq 1} \sum_{n_1, ..., n_k} \frac {a(n_1, ..., n_k) k!} {{m \choose n_1, ..., n_k} {m - 1 \choose k - 1}} = \sum_{k \geq 1} \frac 1 {{m - 1 \choose k - 1}} \sum_{n_1, ..., n_k} \frac {a(n_1, ..., n_k)} {\frac 1 {k!} {m \choose n_1, ..., n_k}}
\end{equation*}
is very similar to the LYM inequality.
The only difference is that elements elements corresponding to different class sizes $n_1, ..., n_k$ might receive slightly different weight, depending on how many duplicate $n_i$ occur in that sequence.

Finally, there is still one interesting thought I had while thinking about the above inequality.
Note that the decomposition
\begin{equation*}
    \Pi_m = \bigcup_{k \geq 1, \ n_1 \leq ... \leq n_k} \{ \pi \in \Pi_m \ | \ |\pi| = k, |\{ S \in \pi \ | \ |S| = j \}| = n_j \}
\end{equation*}
corresponds exactly to the core $\mathcal{C}(\Pi_m)$ of $\Pi_m$ as discussed above.
In particular, the decomposition is exactly
\begin{equation*}
    \Pi_m = \bigcup \ \{ \mathrm{Aut}(\Pi_m).\pi \ | \ \pi \in \Pi_m \}
\end{equation*}

\subsection*{Conclusion}
In this work, we have seen how to generalize the decomposition into layers of $\powerset(n)$ to other posets, which we call graded.
Based on this, we could state an analogue of the LYM inequality for those posets, and give a characterization of it.
Further, we found that indeed, the inequality does not hold for arbitrary graded posets, so the posets with LYM-inequality form a proper subclass of all graded posets.
We have both seen an example of a graded poset with LYM-inequality (that is not $\powerset(n)$) and an example of a graded poset without LYM-inequality.

Additionally, during the attempts to prove the first of those statements, we discovered a polynomial time algorithm (in the size of the poset) that can check whether a given graded poset has an LYM inequality.
By considering the situation modulo symmetry, we could improve the algorithm such that its running time is polynomial in the count of orbits of the poset under its automorphism group.
Of course, this improvement requires knowledge of the automorphism group, but in many cases (like for partitions or multisets), the elements of the poset consist of many ``parts'' and the automorphism group is given by all the permutations of those parts.
We have implemented an ran this algorithm for the case of multisets.

Finally, we had a short look at the literature, and included the proof that the collection of multisets indeed has an LYM-inequality.
Interestingly, this topic of LYM generalizations seems to have been an important research topic in the last century.

There are some questions that arise naturally from this work.
Primarily, it might be the case that there are inequalities similar to the LYM inequality, as e.g. in Prop.~\ref{prop:lym_variant_partitions}.
To me, it looks like one could at least provide an LYM-variant for arbitrary graded posets by weighting elements in each orbit of the automorphism group differently.
The idea would then be similar to Prop.~\ref{prop:core_of_multisets}, but yield another result, without requiring the relatively complicated assumption.

Of course, there are quite a lot of common graded posets, and it is an interesting question which of those satisfy the LYM inequality.
I have seen quite a lot of literature about that, so it would certainly be a promising way to continue this work.

In the end, this Miniproject generalized some aspects of the first section of the lecture to a wider class of posets.

\section{Appendix (Code described in Part VI)}
To find the computational results mentioned above, I have implemented the described algorithm in Rust.
For this, I use my own library containing a collection of math functions as well as Mosek LP solver \cite{mosek} and its rust binding library.
After installing the Mosek binaries manually, it therefore suffices to add
\begin{lstlisting}
    feanor_la = { 
        git = "https://github.com/feanorTheElf/FeanorLA.git"  
    }
    mosek = "0.2"
\end{lstlisting}
to the dependencies of your \texttt{Cargo.toml} file.

After all dependencies are present, the following code will yield the result.
\lstinputlisting[language = rust]{./src/main.rs}

\printbibliography
\end{document}